---
title: "AsianFoodAccess_R"
author: "Rujia Xie"
date: "`r Sys.Date()`"
output: html_document
---

# Load Packages and Dataset

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(corrplot)
library(tidyverse)
library(tidymodels)
library(readxl)
library(car)
library(knitr)
library(caret)
library(leaps)
library(MASS)
library(readxl)
library(prettyR)
options(digits = 4)
options(scipen=4)
#Set working dir to this dir
setwd(getwd())

#Read Fulldata from /Data
FullDataset <- read_excel("../Data/FullDataset_Input.xlsx", 
    col_types = c("numeric", "numeric", "numeric", "numeric", "text", 
                  "numeric", "numeric", "numeric", "numeric", "numeric",
                  "numeric", "numeric","numeric","numeric","numeric","numeric", 
                  "numeric", "numeric", "numeric"))
```


# Descriptive Statistics of Variables

## Descriptive Statistics of the Dependent Variable and Covariates

```{r descriptives of the Dependent Variable and Covariates}
#Mean and standard deviation of Asian American and Pacific Islander (AAPI) Diabetes Rate of all San Diego Census Tracts (CTs)
#Unit: number of hospitalizations and emergency department discharges for diabetes per 100,000 AAPI population per year
FullDataset %>%
  summarize(mean(Diabetes_Rate),
            sd(Diabetes_Rate))%>%
  kable()

#Mean and standard deviation of Percent AAPI Population of all San Diego CTs
FullDataset %>%
  filter(Percent_AAPI != "NA") %>%
  summarize(mean(Percent_AAPI),
            sd(Percent_AAPI)) %>%
  kable()

#Mean and standard deviation of Percent below 200% Poverty Line of all San Diego CTs
FullDataset %>%
  filter(Percent_Below_200_Poverty != "NA") %>%
  summarize(mean(Percent_Below_200_Poverty),
            sd(Percent_Below_200_Poverty))%>%
  kable()

#Mean and standard deviation of Percent of Population over 65 of all San Diego CTs
FullDataset %>%
  filter(Over65Percent != "NA") %>%
  summarize(mean(Over65Percent),
            sd(Over65Percent))%>%
  kable()

#Mean and standard deviation of Percent of Population who are obese of all San Diego CTs
FullDataset %>%
  filter(Obesity_Percent != "NA") %>%
  summarize(mean(Obesity_Percent),
            sd(Obesity_Percent))%>%
  kable()

#Mean and standard deviation of Population Density of all San Diego CTs
#Unit: number of people per square mile
FullDataset %>%
  filter(Density2019 != "NA") %>%
  summarize(mean(Density2019),
            sd(Density2019))%>%
  kable()
```

## Descriptive Statistics of Independent (Food Access) Variables

```{r descriptives of Food Access Variables}
#Mean and standard deviation of percent of CT population >½ mile to the nearest supermarket of the 564 out of 626 San Diego CTs with data available
FullDataset %>%
  filter(lapophalfshare != "NA") %>%
  summarize(mean(lapophalfshare),
            sd(lapophalfshare))%>%
  kable()

#Mean and standard deviation of percent of CT’s AAPI population >½ mile to the nearest supermarket of the 564 out of 626 San Diego CTs with data available
FullDataset %>%
  filter(laaapihalfshare != "NA") %>%
  summarize(mean(laaapihalfshare),
            sd(laaapihalfshare))%>%
  kable()

#Mean and standard deviation of the Yelp ratio of Asian restaurants to non-Asian restaurants of all San Diego CTs 
FullDataset %>%
  filter(ratio_AR_2_NAR_adj != "NA") %>%
  summarize(mean(ratio_AR_2_NAR_adj),
            sd(ratio_AR_2_NAR_adj))%>%
  kable()

#Mean and standard deviation of the Yelp ratio of Asian groceries to non-Asian groceries of all San Diego CTs 
FullDataset %>%
  filter(ratio_AG_2_NAG_adj != "NA") %>%
  summarize(mean(ratio_AG_2_NAG_adj),
            sd(ratio_AG_2_NAG_adj))%>%
  kable()

#Mean and standard deviation of the Yelp ratio of Asian restaurants to AAPI population of all San Diego CTs 
FullDataset %>%
  filter(ratio_AR_2_AAPI != "NA") %>%
  summarize(mean(ratio_AR_2_AAPI),
            sd(ratio_AR_2_AAPI))%>%
  kable()

#Mean and standard deviation of the Yelp ratio of Asian groceries to AAPI population of all San Diego CTs 
FullDataset %>%
  filter(ratio_AG_2_AAPI != "NA") %>%
  summarize(mean(ratio_AG_2_AAPI),
            sd(ratio_AG_2_AAPI))%>%
  kable()

#Mean and standard deviation of the Yelp proportion of CT Asian restaurants to the total number of Asian restaurants in San Diego of all San Diego CTs
FullDataset %>%
  filter(proportion_ctAR_2_sdAR != "NA") %>%
  summarize(mean(proportion_ctAR_2_sdAR),
            sd(proportion_ctAR_2_sdAR))%>%
  kable()

#Mean and standard deviation of the Yelp proportion of CT Asian groceries to the total number of Asian groceries in San Diego of all San Diego CTs
FullDataset %>%
  filter(proportion_ctAG_2_sdAG != "NA") %>%
  summarize(mean(proportion_ctAG_2_sdAG),
            sd(proportion_ctAG_2_sdAG))%>%
  kable()
```

# Correlation Matrices of Variables

Correlation matrices were calculated to examine the correlation among the covariates
(first matrix) and among the food access variables (second matrix)

```{r correlation-matrix-1, fig.height=8, fig.width=12}
#A subset of variables were selected from the FullDataset to run a correlation matrix among the covariates
CovariateSubset <- FullDataset [c(2,3, 11:13)] %>%
  rename(FARA_AAPI = laaapihalfshare,
         Population_Density = Density2019,
         Percent_Obesity = Obesity_Percent,
         Percent_Over_65 = Over65Percent)
correlation_covariate<-cor(CovariateSubset, method="spearman", 
                           use="pairwise.complete.obs")

#In the correlation matrix, a correlation coefficient was shown in each cell.
#A red filling in the cell indicated a positive correlation, and a blue filling indicated a negative correlation. 
#The P value for each correlation was also calculated.Correlations with p>.05 were crossed out in the matrix.
pmatrix <- cor.mtest(CovariateSubset, conf.level = .95)
corrplot(correlation_covariate, p.mat = pmatrix$p, sig.level = .05,
         insig = "pch", pch.cex = 3, pch.col = "red", type = "lower",
         method="color", addCoef.col = "black", number.cex = 1.2, 
         tl.cex = 1.2, tl.srt = 45, tl.col = "black")
```

```{r correlation-matrix-2, fig.height=8, fig.width=12}
#A subset of variables were selected from the FullDataset to run a correlation matrix among the food access variables
FoodAccessSubset <- FullDataset [c(2:4, 14:19)] %>%
  rename(FARA_AAPI = laaapihalfshare,
         FARA_CT = lapophalfshare,
         AR_NAR = ratio_AR_2_NAR_adj,
         AG_NAG = ratio_AG_2_NAG_adj,
         AR_AAPI = ratio_AR_2_AAPI,
         AG_AAPI = ratio_AG_2_AAPI,
         ctAR_sdAR = proportion_ctAR_2_sdAR,
         ctAG_sdAG = proportion_ctAG_2_sdAG)
correlation_foodaccess<-cor(FoodAccessSubset, method="spearman", 
                            use="pairwise.complete.obs")

#In the correlation matrix, a correlation coefficient was shown in each cell.
#A red filling in the cell indicated a positive correlation, and a blue filling indicated a negative correlation. 
#The P value for each correlation was also calculated (shown in the p matrix).Correlations with p>.05 were crossed out in the matrix.
pmatrix <- cor.mtest(FoodAccessSubset, conf.level = .95)
corrplot(correlation_foodaccess, p.mat = pmatrix$p, sig.level = .05,
         insig = "pch", pch.cex = 3, pch.col = "red", type = "lower",
         method="color", addCoef.col = "black", number.cex = 1, 
         tl.cex = 1.2, tl.srt = 45, tl.col = "black",)

print(pmatrix)
```
As shown in the second matrix, FARA_CT and FARA_AAPI were strongly correlated 
with each other with Spearman’s correlation coefficient ρ = 0.98 at p<.001. 
In addition, AR_AAPI and ctAR_sdAR were strongly correlated with ρ=.91 (p<.001).
Thus, FARA_CT and ctAR_sdAR were dropped from further analyses.


# Ordinary Least Squares Regression

After dropping FARA_CT and ctAR_sdAR, the modeling focused on the six remaining
variables: FARA_AAPI, AR_NAR, AR_AAPI, AG_NAG, and AG_AAPI, and ctAG_sdAG.
All models used the scale() function to calculate standardized coefficients. In
all six OLS models, the covariates are population density, percent of
population over 65, and percent of population who are obese.

## 1. OLS Model with FARA_AAPI

```{r FARA_AAPI OLS}
#Food Access Variable: percent of CT’s AAPI population >½ mile to the nearest supermarket
summary(glmFARA_AAPI <- glm(Diabetes_Rate ~ scale(laaapihalfshare) +
                   scale(Density2019) + scale(Over65Percent) + 
                     scale(Obesity_Percent),
        data = FullDataset, na.action = na.exclude))
confint(glmFARA_AAPI)

#Residuals of the OLS model were exported to a new column in FullDataset, "FARA_AAPI"
FullDataset$FARA_AAPI <- resid(glmFARA_AAPI, "pearson")
```

## 2. OLS Model with AR_NAR

```{r AR_NAR OLS}
#Food Access Variable: the Yelp ratio of CT Asian restaurants to CT non-Asian restaurants
summary(glmAR_NAR <- glm(Diabetes_Rate ~ scale(ratio_AR_2_NAR_adj) +
                       scale(Density2019) + scale(Over65Percent) + 
                         scale(Obesity_Percent),
        data = FullDataset, na.action = na.exclude))
confint(glmAR_NAR)

#Residuals of the OLS model were exported to a new column in FullDataset, "AR_NAR"
FullDataset$AR_NAR <- resid(glmAR_NAR, "pearson")
```

## 3. OLS Model with AG_NAG

```{r AG_NAG OLS}
#Food Access Variable: the Yelp ratio of CT Asian groceries to CT non-Asian groceries
summary(glmAG_NAG <- glm(Diabetes_Rate ~ scale(ratio_AG_2_NAG_adj)
                  + scale(Density2019) + scale(Over65Percent) + 
                    scale(Obesity_Percent),
        data = FullDataset, na.action = na.exclude))
confint(glmAG_NAG)

#Residuals of the OLS model were exported to a new column in FullDataset, "AG_NAG"
FullDataset$AG_NAG <- resid(glmAG_NAG, "pearson")
```

## 4. OLS Model with AR_AAPI

```{r AR_AAPI OLS}
#Food Access Variable: the Yelp ratio of CT Asian restaurants to CT AAPI population
summary(glmAR_AAPI <- glm(Diabetes_Rate ~ scale(ratio_AR_2_AAPI) + 
                            scale(Density2019) + scale(Over65Percent) + 
                            scale(Obesity_Percent),
        data = FullDataset, na.action = na.exclude))
confint(glmAR_AAPI)

#Residuals of the OLS model were exported to a new column in FullDataset, "AR_AAPI"
FullDataset$AR_AAPI <- resid(glmAR_AAPI, "pearson")
```

## 5. OLS Model with AG_AAPI

```{r AG_AAPI OLS}
#Food Access Variable: the Yelp ratio of CT Asian groceries to CT AAPI population
summary(glmAG_AAPI <- glm(Diabetes_Rate ~ scale(ratio_AG_2_AAPI) + 
                            scale(Density2019) + scale(Over65Percent) + 
                            scale(Obesity_Percent),
        data = FullDataset, na.action = na.exclude))
confint(glmAG_AAPI)

#Residuals of the OLS model were exported to a new column in FullDataset, "AG_AAPI"
FullDataset$AG_AAPI <- resid(glmAG_AAPI, "pearson")
```

## 6. OLS Model with ctAG_sdAG

```{r ctAG_sdAG OLS}
#Food Access Variable: the Yelp proportion of CT Asian grocery stores to the total number of Asian grocery stores in San Diego County
summary(glmctAG_sdAG <- glm(Diabetes_Rate ~ scale(proportion_ctAG_2_sdAG) + 
                              scale(Density2019) + scale(Over65Percent) + 
                              scale(Obesity_Percent),
        data = FullDataset, na.action = na.exclude))
confint(glmctAG_sdAG)

#Residuals of the OLS model were exported to a new column in FullDataset, "ctAG_sdAG"
FullDataset$ctAG_sdAG <- resid(glmctAG_sdAG, "pearson")
```

# Export the New FullDataset with OLS Residuals as "FullDataset_with_OLS.xlsx"

To test the spatial autocorrelation of OLS residuals, the new FullDataset with
OLS residual columns was exported as "FullDataset_with_OLS.xlsx" and imported
into ArcGIS for spatial autocorrelation analyses.

```{r}
library(rio)
#Note: Un-comment the below line and run install_formats() if there is issue exporting the data to .xlsx
# install_formats()
export(FullDataset, "../Data/FullDataset_with_OLS.xlsx")
```

# Spatial Lag Regression

To prepare the dataset for spatial lag regression, the "FullDataset_with_OLS.xlsx" file 
was first imported to ArcGIS to generate a shapefile (SpatialDataset.*) for the spatial lag 
regression. The shapefile was then exported from ArcGIS and imported into R as 
"SpatialDataset"

```{r data-preparation}
library(sf)
library(spatialreg)
library(spdep)
SpatialDataset <- st_read("../Data/SpatialDataset/SpatialDataset.shp")

# Building spatial weights using Queen contiguity
SpatialDatasetnb <- poly2nb(SpatialDataset, queen = TRUE)
SpatialDatasetW <- nb2listw(SpatialDatasetnb)
```

The spatial lag regression was then run for each of the six food access
variables. In each model, "Diabetes_R" was diabetes rate, Density201" was 
population density, "Over65Perc" was percent over 65, "Obesity_Pe" was 
obesity percentage. Like OLS models, all SLR models used the scale() function 
to calculate standardized coefficients. 

## 1. SLR Model with FARA_AAPI

```{r FARA_AAPI SLR}
#Food Access Variable: percent of CT’s AAPI population >½ mile to the nearest supermarket
summary(FARA_AAPI_SLR <- lagsarlm(Diabetes_R ~ scale(laaapihalf) + 
                                    scale(Density201) + scale(Over65Perc) + 
                                    scale(Obesity_Pe),
                          data = SpatialDataset, na.action = na.exclude, 
                          listw = SpatialDatasetW))

#Residuals of the SLR model were exported to a new column in FullDataset, "FARA_AAPI_SLR"
FullDataset$FARA_AAPI_SLR <- resid(FARA_AAPI_SLR, "pearson")
```

## 2. SLR Model with AR_NAR

```{r AR_NAR SLR}
#Food Access Variable: the Yelp ratio of CT Asian restaurants to CT non-Asian restaurants
summary(AR_NAR_SLR <- lagsarlm(Diabetes_R ~ scale(ratio_AR_2) + 
                                 scale(Density201) + scale(Over65Perc) + 
                                 scale(Obesity_Pe),
                          data = SpatialDataset, na.action = na.exclude, 
                          listw = SpatialDatasetW))

#Residuals of the SLR model were exported to a new column in FullDataset, "AR_NAR_SLR"
FullDataset$AR_NAR_SLR <- resid(AR_NAR_SLR, "pearson")
```

## 3. SLR Model with AG_NAG

```{r AG_NAG SLR}
#Food Access Variable: the Yelp ratio of CT Asian groceries to CT non-Asian groceries
summary(AG_NAG_SLR <- lagsarlm(Diabetes_R ~ scale(ratio_AG_2) + 
                                 scale(Density201) + scale(Over65Perc) + 
                                 scale(Obesity_Pe),
                          data = SpatialDataset, na.action = na.exclude, 
                          listw = SpatialDatasetW))

#Residuals of the SLR model were exported to a new column in FullDataset, "AG_NAG_SLR"
FullDataset$AG_NAG_SLR <- resid(AG_NAG_SLR, "pearson")
```

## 4. SLR Model with AR_AAPI

```{r AR_AAPI SLR}
#Food Access Variable: the Yelp ratio of CT Asian restaurants to CT AAPI population
summary(AR_AAPI_SLR <- lagsarlm(Diabetes_R ~ scale(ratio_AR_3) + 
                                  scale(Density201) + scale(Over65Perc) + 
                                  scale(Obesity_Pe),
                          data = SpatialDataset, na.action = na.exclude, 
                          listw = SpatialDatasetW))

#Residuals of the SLR model were exported to a new column in FullDataset, "AR_AAPI_SLR"
FullDataset$AR_AAPI_SLR <- resid(AR_AAPI_SLR, "pearson")
```

## 5. SLR Model with AG_AAPI

```{r AG_AAPI SLR}
#Food Access Variable: the Yelp ratio of CT Asian groceries to CT AAPI population
summary(AG_AAPI_SLR <- lagsarlm(Diabetes_R ~ scale(ratio_AG_3) + 
                                  scale(Density201) + scale(Over65Perc) + 
                                  scale(Obesity_Pe),
                          data = SpatialDataset, na.action = na.exclude, 
                          listw = SpatialDatasetW))

#Residuals of the SLR model were exported to a new column in FullDataset, "AG_AAPI_SLR"
FullDataset$AG_AAPI_SLR <- resid(AG_AAPI_SLR, "pearson")
```

## 6. SLR Model with ctAG_sdAG

```{r ctAG_sdAG SLR}
#Food Access Variable: the Yelp proportion of CT Asian grocery stores to the total number of Asian grocery stores in San Diego County
summary(ctAG_sdAG_SLR <- lagsarlm(Diabetes_R ~ scale(proporti_1) + 
                                    scale(Density201) + scale(Over65Perc) + 
                                    scale(Obesity_Pe),
                          data = SpatialDataset, na.action = na.exclude, 
                          listw = SpatialDatasetW))

#Residuals of the SLR model were exported to a new column in FullDataset, "ctAG_sdAG_SLR"
FullDataset$ctAG_sdAG_SLR <- resid(ctAG_sdAG_SLR, "pearson")

```

# Export the New FullDataset with SLR Residuals as "FullDataset_with_SLR.xlsx"

To test the spatial autocorrelation of SLR residuals, the new FullDataset with
SLR residual columns was exported as "FullDataset_with_SLR.xlsx" and imported
into ArcGIS for spatial autocorrelation analyses.

```{r FullDatasetSpatial}
library(rio)
#Note: Un-comment the below line and run install_formats() if there is issue exporting the data to .xlsx
# install_formats()
export(FullDataset, "../Data/FullDataset_with_SLR.xlsx")
```


